---
title: California Congressional Gerrymandering Analysis
author: Dory Yang
format: 
  dashboard:
    orientation: rows
    theme: yeti
server: shiny

---
# Executive Summary {.sidebar}
### Research Question

hi

"Will the adoption of the new district map (AB 604) lead to an increase in partisan advantage relative to the 2024 Actual Map? If so, by how much?"

### Methodology Overview

The 2024 Congressional General Election results were interpolated onto the proposed AB 604 district boundaries using Area-Weighted Interpolation (Approach A).
Key metrics— Efficiency Gap (EG) and Mean–Median Difference (MM) —were calculated for both the actual and hypothetical maps to enable a direct comparison of partisan advantage.

#  Project Data Setup
## Initial Data & Metric Calculation
```{r}
#| label: data-setup
library(tidyverse)
library(sf)
library(leaflet)
EG_old <- 0.045
MM_old <- 0.021
Seats_DEM_old <- 5
Seats_REP_old <- 5
EG_new <- -0.2820
MM_new <- -0.0011
Seats_DEM_new <- 9
Seats_REP_new <- 1

Metric <- c("Seats Won (DEM)", "Efficiency Gap (EG)", "Mean-Median Diff (MM)")
Old_Map <- c(Seats_DEM_old, EG_old, MM_old)
New_Map <- c(Seats_DEM_new, EG_new, MM_new)
Change <- New_Map - Old_Map

comparison_data <- data.frame(
  Metric = Metric,
  `2024 Actual Map (Old)` = Old_Map,
  `2025 AB 604 Map (New)` = New_Map,
  Change = Change
)

if (!exists("final_results")) {
  final_results <- data.frame(
    New_District_ID = c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10"),
    Total_Allocated_DEM_Votes = c(244923, 254434, 199449, 214025, 233157, 177364, 189037, 214908, 151258, 245113),
    Total_Allocated_REP_Votes = c(216476, 174137, 186079, 163544, 306518, 154599, 143874, 102422, 107600, 126260),
    Total_Votes = c(461399, 428571, 385528, 377570, 539675, 331962, 332911, 317330, 258858, 371373),
    DEM_Vote_Share = c(0.531, 0.594, 0.517, 0.567, 0.432, 0.534, 0.568, 0.677, 0.584, 0.660),
    Winner = c("DEM", "DEM", "DEM", "DEM", "REP", "DEM", "DEM", "DEM", "DEM", "DEM")
  )
}
```

## AB 604 District Map

```{r seat-plot}
#| label: setup-plot
#| fig-width: 7
#| fig-height: 5
seat_comparison <- data.frame(Map = c("2024 Actual", "2025 AB 604"),Seats = c(Seats_DEM_old, Seats_DEM_new))
seat_chart <- ggplot(seat_comparison, aes(x = Map, y = Seats, fill = Map)) + geom_bar(stat = "identity") + geom_text(aes(label = Seats), vjust = -0.5) +labs(title = "Democratic Seats Won: Old vs. New Map", y = "Number of Seats") +
    theme_minimal() 
seat_chart

```


# Comparison Table

```{r}
#| label: metric-table
comparison_data
```

# Policy Interpretation

```{r}
#| label: policy-interpretation
EG_diff <- EG_new - EG_old
MM_diff <- MM_new - MM_old

EG_text <- if (EG_diff < 0) {
  paste0("The Efficiency Gap (EG) strongly favors the DEM Party, shifting from ", round(EG_old*100, 1), "% to ", round(EG_new*100, 1), "%.")
} else if (EG_diff > 0.01) {
  paste0("The Efficiency Gap (EG) favors the REP Party, shifting from ", round(EG_old*100, 1), "% to ", round(EG_new*100, 1), "%.")
} else {
  "The Efficiency Gap change is minimal, suggesting no significant partisan bias shift."
}

MM_text <- if (abs(MM_new) < abs(MM_old) * 0.5) {
  "The Mean-Median Difference indicates a more uniform vote distribution in the new map, minimizing wasted votes."
} else {
  "The Mean-Median Difference indicates a similar level of vote clustering across both maps."
}

print(EG_text)
print(MM_text)
```

# Hypothetical AB 604 (Part 5)

## Row (Seat Summary)

```{r}
#| label: seat-summary
print(paste0("DEM Seats Won: ", Seats_DEM_new))
print(paste0("REP Seats Won: ", Seats_REP_new))
```

## Column (Seat Allocation)
```{r}
#| label: final-results-table
final_results %>%
  mutate(
    DEM_Share = paste0(round(DEM_Vote_Share * 100, 2), "%")
  ) %>%
  select(-DEM_Vote_Share) %>%
  rename(
    District_ID = New_District_ID,
    DEM_Votes = Total_Allocated_DEM_Votes,
    REP_Votes = Total_Allocated_REP_Votes,
    Total_Votes = Total_Votes,
    Winner = Winner
  )
```

## Map Comparison

### Row

#### Column
